Eco-evolutionary dynamics of finite populations from first principles

### 2023-04-08 ind_files/06_nontechnical_discussion.tex
Add a reference

Several specific finite population models illustrate that evolution in finite population can proceed in a direction different from that predicted by the infinite population limit~\citep{parsons_consequences_2010,melbinger_evolutionary_2010, houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,debarre_evolutionary_2016, behar_fluctuations-induced_2016, constable_demographic_2016,veller_drift-induced_2017,abu_awad_effects_2018,parsons_pathogen_2018,mcavoy_public_2018,mcleod_social_2019}. Indeed, in some stochastic models, the outcome of finite population models can be exactly opposite to that of the infinite population limit, a phenomenon sometimes referred to as `reversing the direction of deterministic evolution'~\citep{constable_demographic_2016,mcleod_social_2019, wang_reproductive_2023}. These effects have been recognized to play an important role in specific finite population models of epidemiology~\citep{kogan_two-strain_2014,parsons_pathogen_2018,day_price_2020}, heterogamety~\citep{veller_drift-induced_2017,saunders_sex_2018}, life-history evolution~\citep{gillespie_natural_1974,kuosmanen_turnover_2022}, and social evolution~\citep{houchmandzadeh_selection_2012,houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,constable_demographic_2016,mcleod_social_2019,wang_reproductive_2023}.

>>>>

Several specific finite population models illustrate that evolution in finite population can proceed in a direction different from that predicted by the infinite population limit~\citep{parsons_consequences_2010,melbinger_evolutionary_2010, houchmandzadeh_selection_2012, houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,debarre_evolutionary_2016, behar_fluctuations-induced_2016, constable_demographic_2016,veller_drift-induced_2017,abu_awad_effects_2018,parsons_pathogen_2018,mcavoy_public_2018,mcleod_social_2019}. Indeed, in some stochastic models, the outcome of finite population models can be exactly opposite to that of the infinite population limit, a phenomenon sometimes referred to as `reversing the direction of deterministic evolution'~\citep{constable_demographic_2016,mcleod_social_2019, wang_reproductive_2023}. These effects have been recognized to play an important role in specific finite population models of epidemiology~\citep{kogan_two-strain_2014,humplik_evolutionary_2014,parsons_pathogen_2018,day_price_2020}, heterogamety~\citep{veller_drift-induced_2017,saunders_sex_2018}, life-history evolution~\citep{gillespie_natural_1974,kuosmanen_turnover_2022}, and social evolution~\citep{houchmandzadeh_selection_2012,houchmandzadeh_fluctuation_2015,chotibut_evolutionary_2015,constable_demographic_2016,mcleod_social_2019,wang_reproductive_2023}.
